Programming Language Survey Paper on Haskell

### 2013-04-21 projects/term_paper/paper.tex
more work on paper

Type checking in Haskell is static and is performed at compile time. Haskell uses type-declarations for functions, in which the types of parameters and return values are explicitly defined. Any use of a function in which parameters do not meet the defined function type is flagged by the compiler, and compilation will be aborted. For example, the following function defines

@@

@@

Haskell also has a String data type. Although Strings are represented by double quotations, they are actually stored as a list of characters. Thus, when specificying a data type, String and [Char] are equivalent.

\begin{minted}{haskell}
let i = 2
let x = 3.14
let c = 'c'
let str = ['H','a','s','k','e','l','l','!']
; shorthand for the above
let str2 = "Haskell!"
let t = (3.14, 'pi')
\end{minted}

    \item Type variables: variables that can be any type (used in polymorphic functions)
\begin{minted}{haskell}
lastTwo :: [a] -> [a]
lastTwo a = [a !! ((length a) - 2), a !! ((length a) - 1)]
\end{minted}
\ei

\subsubsection{Type Checking}
\bi
    \item Haskell: static typing, type inferencing by context
\begin{minted}{haskell}
let myStr = "2"
; will result in an exception as the desired type to convert to is
; ambiguous since Haskell has no context to infer the type
let myNum = read myStr
; not ambiguous, as the compiler infers the Int type is desired
let myNum2 = (read myStr) + 2
\end{minted}
    \item Type classes: Eq, Ord, Show/Read, Bounded, Enum, Num/Integral/Floating
\ei

>>>>

@@

\begin{minted}{haskell}
let x = 2
let pi = 3.14
let c = 'c'
\end{minted}

@@

\begin{minted}{haskell}
let evens = [2, 4, 6, 8, 10]	-- list
let t = (3.14, 'pi')		-- tuple
\end{minted}

Haskell also has a String data type. Although Strings are represented by double quotations, they are actually stored as a list of characters. Thus, when specifiying a data type, String is equivalent to [Char].

\begin{minted}{haskell}
let str1 = ['H','a','s','k','e','l','l','!']
let str2 = "Haskell!" 	-- a string representation of str1
\end{minted}

In addition to these basic data types, Haskell also offers type variables. Type variables are variables whose type has not been explicitly declared; rather, they can be any type. By defining a function without declaring data types, the variables in the function can be used as any type and the function can be polymorphic. In the following example, the function can handle a list of any type because its uses a type variable.

\begin{minted}{haskell}
lastTwo :: [a] -> [a]
lastTwo a = [a !! ((length a) - 2), a !! ((length a) - 1)]
\end{minted}

\subsubsection{Type Checking}
Type checking in Haskell is performed statically at compile time. Haskell uses type-declarations for functions, in which the types of parameters and return values are explicitly defined. Any use of a function in which parameters do not meet the defined function type is flagged by the compiler, and compilation will be aborted. In addition, ues type inferencing by context. The type system in Haskell deduces obvious types and makes them concrete. In the following example, although the type of myNum is unknown at compile time, the compiler infers that the Int type is desired because myNum is added to an Int.

\begin{minted}{haskell}
let myStr = "2" 
let myNum2 = (read myStr) + 2
\end{minted}

If the context does not suggest the type, as in the example below, then an exception will result.

\begin{minted}{haskell}
let myStr = "2" 
let myNum = read myStr
\end{minted}

Haskell uses typeclasses to categorize the type system. Typeclasses define the behavior of a general type. A type must support the definition of its typeclass, but may also be more specific. For example, all of the operators in Haskell are categorized under the Eq typeclass. Other examples of typeclasses in Haskell include Ord, Show, Read, Enum, Num, Integral, and Floating. Typeclasses can be used to allow multiple types to be passed to a function. For example, to create a function that will accept both Int and Integer types, it's as simple as declaring its typeclass as Integral.
